---
title: "Pset 4"
author: "Will Smiles"
date: "2/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load all necessary libraries
library(dbplyr)
library(tidyverse)
library(janitor)
library(gt)
library(lubridate)
library(readxl)
```

# Question 1 
```{r question_1, message=FALSE, echo=FALSE,}
# Read in initial data
poll <- read_csv("ps_4_elections-poll-nc09-3.csv")

# Question 1

# Madlib 1 
dem <- poll %>% 
  # Filter for rows where response value is "Dem" 
  filter(response == "Dem") %>% 
  # nrow() calculates number of rows in the tbl, giving us the number of Dem responses
  nrow()

# Madlib 2
rep <- poll %>% 
  # Filter for rows where hrep_fav value is "Favorable" 
  filter(response == "Rep")

undecided <- poll %>% 
  # Filter for rows where response value is "Und" 
  filter(response == "Und")

# Subtract number of rows of undecided tbl from number of rows of rep  tbl to find difference in value
rep_und_diff <- nrow(rep) - nrow(undecided)


# Madlib 3 
gender_diff <- poll %>%
  # Filter for rows where 'gender' value does not equal 'gender_combined' value
  filter(gender != gender_combined) %>% 
   # nrow() calculates number of rows in the tbl, giving us the number of differences
  nrow()

# Madlib 4
white_diff <- poll %>% 
  # Filter for rows where 'race_eth' value equals "White" but "file_race_black" does not equal "White"
  filter(race_eth == "White") %>% 
  filter(file_race_black != "White") %>% 
  # nrow() calculates number of rows in the tbl, giving us the number of differences
  nrow()

# Madlib 5
rep_time <- poll %>% 
  # Filter for rows where response is "Rep"
  filter(response == "Rep") %>% 
  # Sort in ascending order by timestamp value
  arrange(timestamp) %>% 
  # Retrieve first row
  slice(1) %>% 
  # Pull timestamp value from 1st Rep row
  pull(timestamp)

dem_time <- poll %>% 
  # Filter for rows where response is "Dem"
  filter(response == "Dem") %>% 
  # Sort in ascending order by timestamp value
  arrange(timestamp) %>% 
  # Retrieve first row
  slice(1) %>% 
  # Pull timestamp value from 1st Dem row
  pull(timestamp)

# Calculate the difference rounded by minute by using minute() function for each pulled timestamp value and subtracting
min_diff <- minute(rep_time) - minute(dem_time)
```

There were `r dem` respondents who supported the Democratic candidate.

There were `r rep_und_diff` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (`gender` and `gender_combined`). There are `r gender_diff` individuals for whom these variables have different values.

There are `r white_diff` respondents listed as “White” under `race_eth` who are not listed as “White” under `file_race_black`.

The first response of Rep came `r min_diff` minutes (rounded to the nearest minute) before the first response of Dem.

# Question 2
```{r question_2, echo=FALSE, results="asis"}
# Quesiton 2
table <- poll %>% 
  # Select for response, final_weight, and race_eth columns
  select(response, final_weight, race_eth) %>% 
  # Group by race_eth and response
  group_by(race_eth, response) %>% 
  # Calculate the total sums of the weighted values
  summarize(total = sum(final_weight)) %>%   
  # Filter for rows without refuse/don't know responses
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  # Set key variable of table spread to response (columns), and set fill= 0 to set all N/A values to 0
  spread(key =  response, value = total, fill=0) %>% 
  # Create all variable that is the sum of all possible response values
  mutate(all = Dem + Rep + Und + `3`) %>% 
  # Alter Dem variable to be percentage of Dem responses (Dem divided by All)
  mutate(Dem = Dem / all) %>% 
  # Alter Rep variable to be percentage of Rep responses (Rep divided by All)
  mutate(Rep = Rep / all) %>% 
  # Alter Und variable to be percentage of Und responses (Und divided by All)
  mutate(Und = Und / all) %>% 
  # Deselect all variable and the 3rd party variable
  select(-all, -`3`) %>% 
  arrange(factor(race_eth, levels = c("White","Black","Hispanic","Asian","Other"))) %>% 
  ungroup() %>% 

# Create gt table object
gt() %>% 
  # Use tab_header function to add title to the table
  tab_header(
    title = "Polling Results in North Carolina 9th Congressional District") %>%
  # Use tab_source_note function to add source caption to the table
  tab_source_note(
    source_note="New York Times Upshot/Siena College 2018 live polls") %>% 
  
  # # Center the values in the table using tab_style function and setting text_align

  
  # Use cols_label function to add labels to each column
  cols_label(
    race_eth = "Race",
    Dem = "Democrat",
    Rep = "Republican",
    Und = "Undecided"
  ) %>%
  

  
  # Use fmt_percent function to format numerical values of Dem, Rep, and Und to a percent
  fmt_percent(columns = vars(Dem, Rep, Und),
              # Round to the nearest percent using decimals argument
              decimals = 0) %>% 
  # Set all zero values in the table to N/A using na_if function
  na_if(0) %>% 
  # Use fmt_missing function to specifically set zero value to "-" in Row 4 of Undecided Column
  fmt_missing(columns = vars(Und), rows = 4) %>% 

  # Take the table file, turn it into html, and send it to our md file
  as_raw_html() %>% as.character() %>% cat() 
```

# Question 3
```{r question_3, echo=FALSE}

violin <- poll %>%
  # Select final_weight and education columns
  select(final_weight, educ) %>% 
  # Filter for rows that do not equal "[DO NOT READ] Refused"
  filter(educ != "[DO NOT READ] Refused")

# Reorder the educ values in the proper order, use fct_rev to reverse the order on the graph because of the later coord_flip()
violin$educ <- fct_rev(factor(violin$educ, levels = c("Graduate or Professional Degree", "Bachelors' degree", "Some college or trade school", "High school", "Grade school")))

# Plot the violin data setting x value to educ and y value to final_weight
ggplot(violin, aes(x=educ, y= final_weight)) +

# Add Violin plot
geom_violin() +
# Add Jitter plot 
geom_jitter(alpha = .4, width = .2) +

# Fip the x and y axes using coord_flip function
coord_flip() +

# Add the appropriate axes titles, subtitles, and caption 
labs(title ="More Educated Matter Less in North Carolina 9th", subtitle ="Poll gives more weight for people who are less likely to participate in polls", x = NULL, y = "Weight Given to Respondent in Calculating Poll Results", caption="New York Times Upshot/Siena College 2018 live polls" )
```


# Question 4
```{r question_4, echo=FALSE}
# Create a barchart for party response percentages according to different age groups
response_by_ager <- poll %>% 
  # Filter out refused age responses
  filter(ager != "[DO NOT READ] Refused") %>% 
  # Filter out 3rd party responses (we did not pay attention to these responses earlier in our table, as they are perhaps unimportant for polling purposes)
  filter(response != 3) %>%
  # Group by ager and response variables
  group_by(ager , response) %>% 
  # Count number of different party response types by age group
  summarize(N = n()) %>% 
  # Create frequency ratio and percentage variables by dividing number of frequency of response types by total number of responses
  mutate(freq = N / sum(N), pct = round((freq*100), 0))

# Create a  bar chart that illustrates the percentage of party responses by age group 
# Set x axis to reflect the age groups, y axis to reflect the percentages, and fill the color according to the response type
ggplot(response_by_ager, aes(x = ager, y = pct, fill = response)) + 
  # Geom_col(position="dodge 2") creates vertical bar chart
  geom_col(position = "dodge2") + labs(x = "Age Group",y = "Percent of Responses", fill = "Response")+
  labs(title="Party Response Percentage By Age Group") +
  
  # Fill apropriate colors for each response party
  scale_fill_manual(values = c("blue", "red", "grey")) +
  
  # Place legend at the top and center the title using plot.title() and element_text functions by setting hjust =.5
  theme(legend.position = "top", plot.title = element_text(hjust = 0.5))
```


